Regularising Lotka--Volterra interaction matrices by species grouping: optimal groupings are unpredictable

### 2022-03-20 article.rnw
Add a comma for ease of reading.
Fix a double space between two words.
Remove a non-essential sentence.

Grouping was performed by constraining appropriate competition coefficients to be equal, reducing the number of free parameters in the model.
Each model fit was performed with one grouping of the interaction matrix's rows and a second, independent grouping of the matrix's columns.
A grouping, known as a partition in mathematics, divides the species into groups such that each species belongs to one and only one group, and such that every group contains at least one species.
Given $n$ species, there may be any integer number of groups from $1$ to $n$.
With $n$ groups, the grouping places each species into its own group, such that all species interact in a unique fashion.
With $1$ group, every species is placed in the same group, such that all are equal in their competitive effects or responses, reminiscent of the neutral model.
Intermediate numbers of groups give more freedom in the groupings, such that there are many possible groupings for any intermediate number of groups. % DBS: Is this sentence essential?
The total number of groupings for $n$ species is given by the Bell numbers, denoted $B_n$ \parencite{bell1934exponential}.
With six species and independently selected row and column groupings, the total number of possible models is ${\left(B_6\right)}^2 = \num{\Sexpr{numbers::bell(6)^2}}$.

@@

The Lotka--Volterra model is not linear in the parameters.
With the rows of the interaction matrix ungrouped, the regression could be reparameterised to be linear in the parameters by treating $r_i \alpha_{ij}$ as a single regression parameter for each $i,j$ pair.
With the rows grouped, however, this becomes impossible since each species still has an independent growth rate.
As such, the models were fitted by the Gauss--Newton method for non-linear least squares.
%The models were fitted to the data using maximum likelihood regression techniques.
%Fitting it using precisely a Lotka--Volterra model prevented the use of a link function, rendering most generalised regression techniques infeasible.
I assumed a normal error distribution of the response variable and to prevent heteroskedasticity and improve normality, I transformed total population growth to \emph{per capita} population growth, which corresponds to the following model:
\[
	\frac{N_{i,t+1} - N_{i,t}}{N_{i,t}} = r_i \left(1 - \sum_j \alpha_{ij} N_{j,t}\right)
.\]

>>>>

Grouping was performed by constraining appropriate competition coefficients to be equal, reducing the number of free parameters in the model.
Each model fit was performed with one grouping of the interaction matrix's rows and a second, independent grouping of the matrix's columns.
A grouping, known as a partition in mathematics, divides the species into groups such that each species belongs to one and only one group, and such that every group contains at least one species.
Given $n$ species, there may be any integer number of groups from $1$ to $n$.
With $n$ groups, the grouping places each species into its own group, such that all species interact in a unique fashion.
With $1$ group, every species is placed in the same group, such that all are equal in their competitive effects or responses, reminiscent of the neutral model.
The total number of groupings for $n$ species is given by the Bell numbers, denoted $B_n$ \parencite{bell1934exponential}.
With six species and independently selected row and column groupings, the total number of possible models is ${\left(B_6\right)}^2 = \num{\Sexpr{numbers::bell(6)^2}}$.

@@

The Lotka--Volterra model is not linear in the parameters.
With the rows of the interaction matrix ungrouped, the regression could be reparameterised to be linear in the parameters by treating $r_i \alpha_{ij}$ as a single regression parameter for each $i,j$ pair.
With the rows grouped, however, this becomes impossible, since each species still has an independent growth rate.
As such, the models were fitted by the Gauss--Newton method for non-linear least squares.
%The models were fitted to the data using maximum likelihood regression techniques.
%Fitting it using precisely a Lotka--Volterra model prevented the use of a link function, rendering most generalised regression techniques infeasible.
I assumed a normal error distribution of the response variable and to prevent heteroskedasticity and improve normality, I transformed total population growth to \emph{per capita} population growth, which corresponds to the following model:
\[
	\frac{N_{i,t+1} - N_{i,t}}{N_{i,t}} = r_i \left(1 - \sum_j \alpha_{ij} N_{j,t}\right)
.\]
